---
title: "Homework 11"
author: "Diana Hackenburg"
date: "April 4, 2018"
output: html_document
---

```{r}
# Preliminaries
library(ggplot2)
library(TeachingDemos)
char2seed("Cruel April")
```

########################################################
# FUNCTION: readData
# read in or generate data frame 
# input: file name (or nothing for demo) 
# output: 3-column data frame of observed data (ID, x, y)
#------------------------------------------------------
```{r}
readData <- function(z=NULL) {
  if(is.null(z)) {
    xVar <- sample(1:5,100,replace=TRUE)
    yVar <- sample(1:5,100,replace=TRUE)
    dF <- data.frame(ID=seq_along(xVar), xVar, yVar)
    
    return(dF)
    
  }
}
readData()
```

########################################################
# FUNCTION: getMetric
# calculate metric for randomization test
# input: 3-column data frame for regression
# output: regression slope (single number from randomization test notes)
#------------------------------------------------------
```{r}
getMetric <- function(z=NULL) {
  if(is.null(z)){
    xVar <- sample(1:5,100,replace=TRUE)
    yVar <- sample(1:5,100,replace=TRUE)
    z <- data.frame(ID=seq_along(xVar), xVar, yVar)
  }
  . <- lm(z[,3]~z[,2]) 
  . <- summary(.)
  . <- .$coefficients[2,1]
  slope <- . 
  return(slope) 
}
getMetric()
```

########################################################
# FUNCTION: shuffleData
# randomization data for regression analysis
# input: 3-column data frame (ID, xVar, yVar)
# output: 3-column data frame (ID, xVar, yVar) (just shuffled up!)
#------------------------------------------------------
```{r}
shuffleData <- function(z=NULL) {
  if(is.null(z)){
    xVar <- sample(1:5,100,replace=TRUE)
    yVar <- sample(1:5,100,replace=TRUE)
    z <- data.frame(ID=seq_along(xVar), xVar, yVar)
  }
  z[,3] <- sample(z[,3]) # reshuffling the y values 
  # de-coupling pattern between x and y 
  # default is replace = FALSE so in one swoop shuffles data randomly
  
  return(z)
}
shuffleData()
```

########################################################
# FUNCTION: getPVAL
# calculate p value for observed, simulated data
# input: list of observed metric and vector of simulated metric
# output: lower, upper tail probability vector 
#------------------------------------------------------
```{r}
getPVAL <- function(z=NULL) {
  if(is.null(z)){
    z <- list(xObs=runif(1), xSim=runif(1000)) }
  pLower <- mean(z[[2]]<=z[[1]]) 
  pUpper <- mean(z[[2]]>=z[[1]])
  
  return(c(pL=pLower,pU=pUpper))
}
getPVAL()
```

######################################
# FUNCTION: plotRanTest
# ggplot graph
# input: list of observed metric and vector of simulated metric
# output: ggplot graph
#-------------------------------------
```{r}
plotRanTest <-function(z=NULL){
  if(is.null(z)){
    z <- list(xObs=runif(1),xSim=runif(1000)) }
  dF <- data.frame(ID=seq_along(z[[2]]),simX=z[[2]])
  p1 <- ggplot(data=dF,mapping=aes(x=simX))
  p1 + geom_histogram(mapping=aes(fill=I("goldenrod"),color=I("black"))) + geom_vline(aes(xintercept=z[[1]],color="blue"))
}
plotRanTest()
```

#---------------------------------------------------
# main body of code 
```{r}
nSim <- 1000 # number of simulations 
Xsim <- rep(NA, nSim) # will hold simulated slopes 

dF <- readData()
Xobs <- getMetric(dF)

# loop through to create 1000 simulated slopes for xSim vector
for (i in seq_len(nSim)){
  Xsim[i] <- getMetric(shuffleData(dF))
}

# creating the list structure that we need for final function
slopes <- list(Xobs, Xsim)
getPVAL(slopes)



plotRanTest(slopes)

#--------------
# Real Data!
# Read CSV into R
MyData <- read.csv(file="Homework11data.csv", header=TRUE, sep=",")  
head(MyData)
dF <- data.frame(ID=seq_along(MyData),MyData)
getMetric(dF)
getMetric()
nSim <- 1000 # number of simulations 
Xsim <- rep(NA, nSim) # will hold simulated slopes 

Xobs <- getMetric(dF)

# loop through to create 1000 simulated slopes for xSim vector
for (i in seq_len(nSim)){
  Xsim[i] <- getMetric(shuffleData(dF))
}

# creating the list structure that we need for final function
slopes <- list(Xobs, Xsim)
getPVAL(slopes)



plotRanTest(slopes)
```
