---
title: "Homework 4"
author: "Diana Hackenburg"
date: "February 7, 2018"
output: html_document
---
Solve these problems, some of which are adapted from:\
_Jones, O., R. Maillardet, and A. Robinson. 2009. Scientific Programming and Simulation Using R. CRC Press, Boca Raton._

1.Suppose x = 1.1, a = 2.2, and b = 3.3. Assign each expression to the value of the variable z and print the value stored in z.


```{r}
x<-1.1
a<-2.2
b<-3.3
# a. x^a^b
z<-x^(a^b)
print(z)
# b. (x^a)^b
z<-(x^a)^b
print(z)
# c. 3x^3 + 2x^2 +1
z<- 3*(x^3)+2*(x^2)+1
print(z)
# d. The digit in the second place of z
floor(z %% floor(z)*10)
print(z)            
```

2. Using the rep and seq functions, create the following vectors:

a.(1,2,3,4,5,6,7,8,7,6,5,4,3,2,1)\
b.(1,2,2,3,3,3,4,4,4,4,5,5,5,5,5)\
c.(5,4,4,3,3,3,2,2,2,2,1,1,1,1,1)
```{r}
#a
a<- c(seq(1:8), seq(from=7, to=1))
print(a)
#b
z<-c(1:5)
b<-rep(x=z,times=z)
print(b)
#c
z<-c(seq(from=5, to=1))
print(z)
c<-rep(x=z,times=rev(z))
print(c)

```
3. Create a vector of two random uniform numbers. In a spatial map, these can be interpreted as x and y coordinates that give the location of an individual (such as a marked forest tree in a plot that has been mapped). Using one of R's inverse trigonometry functions (asin(), acos(), or atan()), convert these numbers into polar coordinates. _(If you don't know what polar coordinates are, read about them on the web or in your calculus textbook)._
```{r}
z<-runif(2)
print(z)
# r = (x^2+y^2)^(1/2)
# \theta = arctan(y/x)
r<-(z[1]^2+z[2]^2)^(1/2)
theta<-(atan(z[2]/z[1]))
polarcoords<-c(r,theta)
print(polarcoords)
```
4. Suppose that queue <- c("sheep", "fox", "owl", "ant") and that queue represents the animals that are lined up to enter Noah's Ark, with the sheep at the front of the line. 
```{r}
queue<-c("sheep","fox","owl","ant")
print(queue)
```
Using R expressions, update the queue successively as:
```{r}
#a. the serpent arrives;
length(queue)
queue[5]<-c("serpent")
print(queue)
#b. the sheep enters the ark;
queue<-queue[-1]
print(queue)
#c. the donkey arrives and talks his way to the front of the line;
queue<-c("donkey",queue)
print(queue)
#d. the serpent gets impatient and leaves;
queue<-queue[-5]
print(queue)
#e. the owl gets bored and leaves;
queue<-queue[-3]
print(queue)
#f.the aphid arrives and the ant invites him to cut in line.
queue<-c(queue[-3],"aphid",queue[3])
print(queue)
#e. Finally, determine the position of the aphid in the line.
which(queue == "aphid")
```
5. Use R to create a vector of all of the integers from 1 to 100 that are not divisible by 2, 3, or 7.

```{r}
v<-seq_len(100)
v[!v%%2==0 & !v%%3==0 & !v%%7==0]

```
6. Create a vector z of 1000 random uniform numbers.
```{r}
z<-runif(1000)

# a.create a vector that contains 3 numbers: the proportion of the numbers in z that are less than 0.10, greater than 0.90, and between 0.45 and 0.55.

proportion<-c(mean(z<0.10),mean(z>0.90),mean(z>0.45 & z<0.55))
print(proportion)


# b.Making successive copies of z, transform your vector of uniform numbers in the following ways &  for each case calculate your vector of 3 numbers to get the new proportions.
set.seed(1)  
#log (base 10) of z
zlog<-log10(z)
proportion<-c(mean(zlog<0.10),mean(zlog>0.90),mean(zlog>0.45 & zlog<0.55))
print(proportion)
#z^2
zsquare<-z^2
proportion<-c(mean(zsquare<0.10),mean(zsquare>0.90),mean(zsquare>0.45 & z<0.55))
print(proportion)
#e^z
expz<-exp(z)
proportion<-c(mean(expz<0.10),mean(expz>0.90),mean(expz>0.45 & z<0.55))
print(proportion)
#square root of z
sqrtz<-sqrt(z)
proportion<-c(mean(sqrtz<0.10),mean(sqrtz>0.90),mean(sqrtz>0.45 & z<0.55))
print(proportion)

```