---
title: "Homework 6"
author: "Diana Hackenburg"
date: "February 21, 2018"
output: html_document
---
#Open Libraries
```{r}
library(ggplot2) # for graphics
library(MASS) # for maximum likelihood estimation
```
#Generate my data
```{r}
z<-read.table("MyDataFile_Homework6.csv",header=TRUE,sep=",", stringsAsFactors=FALSE)
str(z)
summary(z)
```
#Plot Histogram of Data
```{r}
p1 <- ggplot(data=z, aes(x=personal, y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) 
print(p1)
```
#Add empirical density curve
```{r}
p1 <-  p1 +  geom_density(linetype="dotted",size=0.75)
print(p1)
```
#Get maximum likelihood parameters for `normal`
```{r}
normPars <- fitdistr(z$personal,"normal")
print(normPars)
str(normPars)
normPars$estimate["mean"] # note structure of getting a named attribute
```
#Plot normal probability density
```{r}
meanML <- normPars$estimate["mean"]
sdML <- normPars$estimate["sd"]
xval <- seq(0,max(z$personal),len=length(z$personal))
stat <- stat_function(aes(x = xval, y = ..y..), fun = dnorm, colour="red", n = length(z$personal), args = list(mean = meanML, sd = sdML))
p1 + stat
``` 
# Plot `exponential` probability density
```{r}
expoPars <- fitdistr(z$personal,"exponential")
rateML <- expoPars$estimate["rate"]
stat2 <- stat_function(aes(x = xval, y = ..y..), fun = dexp, colour="blue", n = length(z$personal), args = list(rate=rateML))
p1 + stat + stat2
```
#Plot `uniform` probability density
```{r}
stat3 <- stat_function(aes(x = xval, y = ..y..), fun = dunif, colour="darkgreen", n = length(z$personal), args = list(min=min(z$personal), max=max(z$personal)))
 p1 + stat + stat2 + stat3
```
#Plot `gamma` probability density
```{r}
gammaPars <- fitdistr(z$personal,"gamma")
shapeML <- gammaPars$estimate["shape"]
rateML <- gammaPars$estimate["rate"]

stat4 <- stat_function(aes(x = xval, y = ..y..), fun = dgamma, colour="goldenrod", n = length(z$personal), args = list(shape=shapeML, rate=rateML))
 p1 + stat + stat2 + stat3 + stat4
```

# Plot `beta` probability density
```{r}
pSpecial <- ggplot(data=z, aes(x=personal/(max(personal + 0.1)), y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) + 
  xlim(c(0,1)) +
  geom_density(size=0.75,linetype="dotted")

betaPars <- fitdistr(x=z$personal/max(z$personal + 0.1),start=list(shape1=1,shape2=2),"beta")
shape1ML <- betaPars$estimate["shape1"]
shape2ML <- betaPars$estimate["shape2"]

statSpecial <- stat_function(aes(x = xval, y = ..y..), fun = dbeta, colour="orchid", n = length(z$personal), args = list(shape1=shape1ML,shape2=shape2ML))
pSpecial + statSpecial
```

#Log Likelihood of Disributions - high log likelihoods are 'best'
```{r}
normPars$loglik
gammaPars$loglik
expoPars$loglik
```

#Gamma wins!

#Simulate Data
```{r}
shapeML
rateML
zSim<-rgamma(n=4545,shape=70.05221,rate=1.910506)
zSimDF<-as.data.frame(zSim)
head(zSimDF)
pSim <- ggplot(data=zSimDF, aes(x=zSim, y=..density..)) + geom_histogram(color="grey60",fill="cornsilk",size=0.2) 
print(pSim)
```

#Simulated data with gamma probability density curve
```{r}
pSim + stat4
```

#Original data with gamma probability density curve
```{r}
p1 <- ggplot(data=z, aes(x=personal, y=..density..)) +
  geom_histogram(color="grey60",fill="cornsilk",size=0.2) 
p1 + stat4
```

##How do the two histogram profiles compare? Do you think the model is doing a good job of simulating realistic data that match your original measurements? Why or why not?
The two histogram profiles are somewhat comparable. The model underestimates the density of values around the mean and doesn't do a good job of simulating the  skewed-right distribution of the original data.
